Derivation Write Up

### 2020-05-28 Documents/ChaoticDynamics/WriteUp.tex
minor mods

\usepackage[margin=2cm]{geometry}
\usepackage{amsmath}
\usepackage{mathtools}

@@

@@

@@

@@

		In order to follow the conventions of spin glass theory for the analysis of disordered systems,
	we rescale the system so that the payoff matrix elements are of order $N^{-1/2}$. The motivation
	for doing this is that, along the way, we will take the limit of the number of actions $N$ to
	infinity. However, doing this will result in numerical underflow of the action probabilities 
	(i.e. the probability of each action goes to zero). To compensate for this, we adjust the
	system so that the sum of the probabilities add to $N$. The scaling goes as follows:

	\begin{subequations}
		\begin{equation}
			a_{ij} = \sqrt{N} \tilde{a_{ij}}
		\end{equation}
		\begin{equation}
			b_{ji} = \sqrt{N} \tilde{b_{ji}}
		\end{equation}
	\end{subequations}

@@

	The need for the factor of $\frac{1}{\sqrt{N}}$ is to follow the conventions of the saddle point
	method of integration. This will become clear after taking the expectation of the generating
	functional. Note that we may drop the notation on time dependence for $x$ and $y$. However, these
	will always be functions of time. Henceforth, we shall not write the tildes on $x, y, a_{ij},
	b_{ij}$. We shall also abbreviate the final term as

	\begin{equation*}
		\rho_{x, i}(t) = \sum_j x_j ln(\frac{x_j}{x_i}) 
	\end{equation*}

@@

	\begin{equation}
	\begin{split}
	\label{eqn::moveddisorder}
		Z(\Vec{\psi}, \Vec{\phi}) = \int D[\Vec{x}, \Vec{\hat{x}}, \Vec{y}, \Vec{\hat{y}}] exp(i
		\sum_i \int dt[\hat{x}_i
		(\frac{\dot{x_i}(t)}{x_i(t)}
			- \tilde{\alpha} \rho_{x, i}(t)+ h_{x, i}(t))]) 
			\\
			\times exp(i \sum_i \int dt[\hat{y}_i
		(\frac{\dot{y_i}(t)}{y_i(t)} 
			- \tilde{\alpha} \rho_{y, i}(t) + h_{y, i}(t))])\\
			\times exp(i \sum_i
		\int dt[x_i(t) \psi_i(t) + y_i(t) \phi_i(t))])\\
		\times exp(i \sum_i \int dt[- \hat{x}_i \alpha \tilde{\tau} \sum_{j} a_
			{ij} 
			y_j +
			\hat{x}_i \tilde{\alpha} \tau \frac{1}{\sqrt{N}} \sum_{i j} x_i a_{ij} y_j - \hat{y}_i
			\alpha 
			\tilde{\tau} \sum_{j} b_
			{ij} 
			x_j +
			\hat{y}_i \tilde{\alpha} \tau \frac{1}{\sqrt{N}} \sum_{i j} y_i b_{ij} x_j]).
	\end{split}
	\end{equation}
	The only difference between (\ref{eqn::generatingfunctional}) and (\ref{eqn::moveddisorder}) is
	that we moved the term containing the quenched disorder into a separate exponential. Since our
	aim is to take an average over all possible realisations of this disorder, we will only need to
	focus on the last exponential which we rewrite as

@@

	It should be noted that we have changed some of the subscripts on the terms. Since these terms
	are all multiplied together and we sum over the subscripts, the letters we choose are of no
	importance and we can exchange them freely. We will define both exponentials in $Q$ as $Q_1$ and
	$Q_2$ respectively.

@@

\begin{align*}
	\frac{\partial f}{\partial \hat{C}_x(t, t')} &\implies C_x(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i <x_i(t) x_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \psi_i(t) \partial \psi_i (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{L}_x(t, t')} &\implies L_x(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i <\hat{x}_i(t) \hat{x}_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial h_{x,i}(t) \partial h_{x, i} (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{K}_x(t, t')} &\implies K_x(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i <x_i(t) \hat{x}_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \Vec{\psi}(t) \partial h_{x, i} (t')}
\bigg|_{\Vec{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
\frac{\partial f}{\partial \hat{C}_y(t, t')} &\implies C_y(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i <y_i(t) y_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \phi_i(t) \partial \phi_i (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{L}_y(t, t')} &\implies L_y(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i <\hat{y}_i(t) \hat{y}_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial h_{y,i}(t) \partial h_{y, i} (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{K}_y(t, t')} &\implies K_y(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i <y_i(t) \hat{y}_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \Vec{\phi}(t) \partial h_{x, i} (t')}
\bigg|_{\Vec{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{A}_{xy}(t, t')} &\implies A_{xy}(t, t') = \lim_{N \rightarrow
	\infty}
N^{-1} \sum_i <\hat{x}_i(t) \hat{y}_i(t')>_\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \partial h_{x, i} (t) \partial h_{y, i} (t')}
\bigg|_{\Vec{\phi} = \Vec{\psi} = \Vec{h} = 0}.\\
\end{align*}

@@

\begin{align}
	G_x(t, t') &= <\frac{\delta x(t)}{\delta \eta_x(t')}> & G_y(t, t') &= <\frac{\delta y(t)}{\delta
	\eta_y(t')}> \\
	<\eta_x(t) \eta_x(t')> &= C_y(t, t') & <\eta_y(t) \eta_y(t')> &= C_x(t, t')\\
	<\mu_x(t) \mu_y(t')> &= C_x(t, t') C_y(t, t') & <\mu_x(t) \mu_x(t')> &= <\mu_y(t) \mu_y(t')> = 0
\end{align}

>>>>

\usepackage[margin=2cm]{geometry}
\usepackage{amsmath,amsfonts}
\usepackage{mathtools}

\usepackage{color}
\newcommand\fb[1]{\textcolor{blue}{FB: #1}}

@@

\fb{the bibfile is missing.}

@@

 \fb{you might discuss in which way any of the points above differs from the relevant reference.}

@@

        \fb{it would be good to have a overall description of the goal of the derivation.}

@@

		In order to follow the conventions of spin glass
                theory for the analysis of disordered systems, we
                rescale the system so that the payoff matrix elements
                are of order $N^{-1/2}$. The motivation for doing this
                is that, along the way, we will take the limit of the
                number of actions $N$ to infinity. However, doing this
                will result in numerical underflow of the action
                probabilities (i.e. the probability of each action
                goes to zero). To compensate for this, we adjust the
                system so that the sum of the probabilities add to
                $N$. The scaling goes as follows:
%
	\begin{subequations}
		\begin{equation}
			a_{ij} = \sqrt{N} \tilde{a_{ij}}
		\end{equation}
		\begin{equation}
			b_{ji} = \sqrt{N} \tilde{b_{ji}}
		\end{equation}
	\end{subequations}

@@

	The need for the factor of $\frac{1}{\sqrt{N}}$ is to follow
        the conventions of the saddle point method of
        integration. This will become clear after taking the
        expectation of the generating functional. Note that we may
        drop the notation on time dependence for $x$ and $y$. However,
        these will always be functions of time. Henceforth, we shall
        not write the tildes on $x, y, a_{ij}, b_{ij}$. We shall also
        abbreviate the final term as
%
	\begin{equation*}
		\rho_{x, i}(t) = \sum_j x_j ln(\frac{x_j}{x_i}) 
	\end{equation*}

@@

	\begin{equation}
	\begin{split}
	\label{eqn::moveddisorder}
		Z(\Vec{\psi}, \Vec{\phi}) = \int D[\Vec{x}, \Vec{\hat{x}}, \Vec{y}, \Vec{\hat{y}}] exp(i
		\sum_i \int dt[\hat{x}_i
		(\frac{\dot{x_i}(t)}{x_i(t)}
			- \tilde{\alpha} \rho_{x, i}(t)+ h_{x, i}(t))]) 
			\\
			\times exp(i \sum_i \int dt[\hat{y}_i
		(\frac{\dot{y_i}(t)}{y_i(t)} 
			- \tilde{\alpha} \rho_{y, i}(t) + h_{y, i}(t))])\\
			\times exp(i \sum_i
		\int dt[x_i(t) \psi_i(t) + y_i(t) \phi_i(t))])\\
		\times exp(i \sum_i \int dt[- \hat{x}_i \alpha \tilde{\tau} \sum_{j} a_
			{ij} 
			y_j +
			\hat{x}_i \tilde{\alpha} \tau \frac{1}{\sqrt{N}} \sum_{i j} x_i a_{ij} y_j - \hat{y}_i
			\alpha 
			\tilde{\tau} \sum_{j} b_
			{ij} 
			x_j +
			\hat{y}_i \tilde{\alpha} \tau \frac{1}{\sqrt{N}} \sum_{i j} y_i b_{ij} x_j]).
	\end{split}
	\end{equation}
	The only difference between (\ref{eqn::generatingfunctional}) and (\ref{eqn::moveddisorder}) is
	that we moved the term \fb{which one exactly?} containing the quenched disorder into a separate exponential. Since our
	aim is to take an average over all possible realisations of this disorder, we will only need to
	focus on the last exponential which we rewrite as

@@

	It should be noted that we have changed some of the subscripts on the terms. Since these terms
	are all multiplied together and we sum over the subscripts, the letters we choose are of no
	importance and we can exchange them freely \fb{perhaps some more justification might be helpful}. We will define both exponentials in $Q$ as $Q_1$ and
	$Q_2$ respectively.

@@

\begin{align*}
	\frac{\partial f}{\partial \hat{C}_x(t, t')} &\implies C_x(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i \langle x_i(t) x_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \psi_i(t) \partial \psi_i (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{L}_x(t, t')} &\implies L_x(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i \langle \hat{x}_i(t) \hat{x}_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial h_{x,i}(t) \partial h_{x, i} (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{K}_x(t, t')} &\implies K_x(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i \langle x_i(t) \hat{x}_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \Vec{\psi}(t) \partial h_{x, i} (t')}
\bigg|_{\Vec{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
\frac{\partial f}{\partial \hat{C}_y(t, t')} &\implies C_y(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i \langle y_i(t) y_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \phi_i(t) \partial \phi_i (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{L}_y(t, t')} &\implies L_y(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i \langle \hat{y}_i(t) \hat{y}_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial h_{y,i}(t) \partial h_{y, i} (t')} \bigg|_{\Vec
{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{K}_y(t, t')} &\implies K_y(t, t') = \lim_{N \rightarrow \infty}
N^{-1} \sum_i \langle y_i(t) \hat{y}_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \Vec{\phi}(t) \partial h_{x, i} (t')}
\bigg|_{\Vec{\phi} = \Vec{\psi} = \Vec{h} = 0}\\
	\frac{\partial f}{\partial \hat{A}_{xy}(t, t')} &\implies A_{xy}(t, t') = \lim_{N \rightarrow
	\infty}
N^{-1} \sum_i \langle \hat{x}_i(t) \hat{y}_i(t')\rangle _\Omega = -\lim_{N \rightarrow \infty} \sum_i
\frac{\partial^2\mathbb{E}[Z(\psi, \phi)]}{\partial \partial h_{x, i} (t) \partial h_{y, i} (t')}
\bigg|_{\Vec{\phi} = \Vec{\psi} = \Vec{h} = 0}.\\
\end{align*}

@@

\begin{align}
	G_x(t, t') &= \langle \frac{\delta x(t)}{\delta \eta_x(t')}\rangle  & G_y(t, t') &= \langle \frac{\delta y(t)}{\delta
	\eta_y(t')}\rangle  \\
	\langle \eta_x(t) \eta_x(t')\rangle  &= C_y(t, t') & \langle \eta_y(t) \eta_y(t')\rangle  &= C_x(t, t')\\
	\langle \mu_x(t) \mu_y(t')\rangle  &= C_x(t, t') C_y(t, t') & \langle \mu_x(t) \mu_x(t')\rangle  &= \langle \mu_y(t) \mu_y(t')\rangle  = 0
\end{align}
